---
title: "CVD Mortality Prediction Analysis"
author: "Sam Fansler"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(survminer)
library(prodlim)
library(pec)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(glmnet)
library(randomForestSRC)
library(ranger)
library(gtsummary)
library(forcats)
library(webshot)
library(formattable)
library(reshape2)
library(nricens)
library(CoxR2)
```

# Read in cleaned data
```{r}

setwd("/Users/samfansler/Library/CloudStorage/OneDrive-JohnsHopkins/Research/Michigan")
data = read.table("data_cvd_mortality.txt", sep = "\t")
train = read.table("train_cvd_mortality.txt", sep = "\t")
test = read.table("test_cvd_mortality.txt", sep = "\t")
metals = read.csv("metal_selected_detection.csv")

```

# Cox Proportional Hazards Model

## Traditional Variables

```{r}

#CVD Death
##Training the model using only the traditional CVD risk factors

CPHM_traditional_cvddeath=coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm, data=train, x=T)

summary(CPHM_traditional_cvddeath)


tbl_regression(CPHM_traditional_cvddeath, exponentiate = T, pvalue_fun = ~style_pvalue(.x, digits = 2),) %>% 
  bold_p(t=0.05) %>%
  modify_footnote(everything() ~ NA, abbreviation = TRUE)

#Survival Curve for survival probability over time
ggsurvplot(survfit(CPHM_traditional_cvddeath), data=train, legend="none", break.y.by=0.05, ylim=c(0.9, 1)) + labs(x = "Time (Years)")


#Calculating C-index
c1=cindex(CPHM_traditional_cvddeath, Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm, data=test)


cindex_CPHM_trad=unname(unlist(c1$AppCindex))


# #All death
# CPHM_traditional_alldeath=coxph(Surv(time, all_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm, data=train, x=T)
# 
# summary(CPHM_traditional_alldeath)
# 
# tbl_regression(CPHM_traditional_alldeath, exponentiate = T, pvalue_fun = ~style_pvalue(.x, digits = 2),) %>% 
#   bold_p(t=0.05) %>%
#   modify_footnote(everything() ~ NA, abbreviation = TRUE)

```
##Adding Blood Metals
```{r}

CPHM_blood_metals_main=coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, x=T)

summary(CPHM_blood_metals_main)

#Calculating C-index
cindex_CPHM_blood_metals_main=cindex(CPHM_blood_metals_main, Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=test)


cindex_CPHM_blood_metals_main=unname(unlist(cindex_CPHM_blood_metals_main$AppCindex))
```

## Including all Metals

```{r}
###Training the model using traditional risk factors + metals

#Just main effects
CPHM_all_metals_main=coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab+ urxucr, data=train, x=T)

summary(CPHM_all_metals_main)

ggsurvplot(survfit(CPHM_all_metals_main), data=train, break.y.by=0.05, ylim=c(0.9, 1))


tbl_regression(CPHM_all_metals_main, exponentiate = T, pvalue_fun = ~style_pvalue(.x, digits = 2),) %>% 
  bold_p(t=0.05) %>%
  modify_footnote(everything() ~ NA, abbreviation = TRUE)

#Calculating C-index
cindex_CPHM_all_metals_main=cindex(CPHM_all_metals_main, Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=test)


cindex_CPHM_all_metals_main=unname(unlist(cindex_CPHM_all_metals_main$AppCindex))


#With squared and interaction terms
CPHM_all_metals_int=coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + bmi + hypertension + dm + urxucs + urxucs^2 + urxumo + urxumo^2 + lbxbpb + lbxbpb^2 + urxutl + urxutl^2 + urxuas + urxuas^2 + urxuco + urxuco^2 + urxuba + urxuba^2 + urxupb + urxupb^2 + urxucd + urxucd^2 + lbxthg + lbxthg^2 + urxuhg + urxuhg^2 + lbxbcd + lbxbcd^2 + urxuur + urxuur^2 + urxudma + urxudma^2 + urxutu + urxutu^2 + urxusb + urxusb^2 + urxuab + urxuab^2 + urxucr + urxucr^2 + (urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr)^2  - lbxbpb:urxucr - lbxthg:urxucr - lbxbcd:urxucr, data=train, x=T)

summary(CPHM_all_metals_int)

#Calculating C-index
cindex_CPHM_all_metals_int=cindex(CPHM_all_metals_int, Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + urxucs + urxucs^2 + urxumo + urxumo^2 + lbxbpb + lbxbpb^2 + urxutl + urxutl^2 + urxuas + urxuas^2 + urxuco + urxuco^2 + urxuba + urxuba^2 + urxupb + urxupb^2 + urxucd + urxucd^2 + lbxthg + lbxthg^2 + urxuhg + urxuhg^2 + lbxbcd + lbxbcd^2 + urxuur + urxuur^2 + urxudma + urxudma^2 + urxutu + urxutu^2 + urxusb + urxusb^2 + urxuab + urxuab^2 + urxucr + urxucr^2 + (urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr)^2 - lbxbpb:urxucr - lbxthg:urxucr - lbxbcd:urxucr, data=test)


cindex_CPHM_all_metals_int=unname(unlist(cindex_CPHM_all_metals_int$AppCindex))
```

## Comparison Table

```{r}
comparison_CPHM=matrix(nrow=4, ncol=1)
comparison_CPHM[1,]=cindex_CPHM_trad
comparison_CPHM[2,]=cindex_CPHM_blood_metals_main
comparison_CPHM[3,]=cindex_CPHM_all_metals_main
comparison_CPHM[4,]=cindex_CPHM_all_metals_int

colnames(comparison_CPHM)=c("C-index")
rownames(comparison_CPHM)=c("Traditional Vars","+Blood Metals (main effects)" ,"+Urinary Metals (main effects)", "+ Metals squared and interaction")

comparison_CPHM
```

## Individual associations with traditional vars

```{r}

cox_age_cvd=coxph(Surv(time, cvd_death) ~ age, data=train, x=T)

summary(cox_age_cvd)

cox_race_cvd=coxph(Surv(time, cvd_death) ~ black + hispanic, data=train, x=T)

summary(cox_race_cvd)

cox_sex_cvd=coxph(Surv(time, cvd_death) ~ male, data=train, x=T)

summary(cox_sex_cvd)

cox_smoke_cvd=coxph(Surv(time, cvd_death) ~ currentsmk, data=train, x=T)

summary(cox_smoke_cvd)

cox_sbp_cvd=coxph(Surv(time, cvd_death) ~ sbp, data=train, x=T)

summary(cox_sbp_cvd)

cox_totchol_cvd=coxph(Surv(time, cvd_death) ~ lbxtc, data=train, x=T)

cox_hdl_cvd=coxph(Surv(time, cvd_death) ~ hdl, data=train, x=T)

summary(cox_hdl_cvd)

cox_bmi_cvd=coxph(Surv(time, cvd_death) ~ overweight + obese, data=train, x=T)

summary(cox_bmi_cvd)

cox_hypertension_cvd=coxph(Surv(time, cvd_death) ~ hypertension, data=train, x=T)

summary(cox_hypertension_cvd)

cox_diabetes_cvd=coxph(Surv(time, cvd_death) ~ dm, data=train, x=T)

summary(cox_diabetes_cvd)
```

## Individual assocations with metals
```{r}
cox_ucs_cvd=coxph(Surv(time, cvd_death) ~ urxucs, data=train, x=T)

summary(cox_ucs_cvd)

cox_umo_cvd=coxph(Surv(time, cvd_death) ~ urxumo, data=train, x=T)

summary(cox_umo_cvd)

cox_bpb_cvd=coxph(Surv(time, cvd_death) ~ lbxbpb, data=train, x=T)

summary(cox_sex_cvd)

cox_utl_cvd=coxph(Surv(time, cvd_death) ~ urxutl, data=train, x=T)

summary(cox_utl_cvd)

cox_uas_cvd=coxph(Surv(time, cvd_death) ~ urxuas, data=train, x=T)

summary(cox_uas_cvd)

cox_uco_cvd=coxph(Surv(time, cvd_death) ~ urxuco, data=train, x=T)

summary(cox_uco_cvd)

cox_uba_cvd=coxph(Surv(time, cvd_death) ~ urxuba, data=train, x=T)

summary(cox_uba_cvd)

cox_upb_cvd=coxph(Surv(time, cvd_death) ~ urxupb, data=train, x=T)

summary(cox_upb_cvd)

cox_ucd_cvd=coxph(Surv(time, cvd_death) ~ urxucd, data=train, x=T)

summary(cox_ucd_cvd)

cox_thg_cvd=coxph(Surv(time, cvd_death) ~ lbxthg, data=train, x=T)

summary(cox_thg_cvd)

cox_uhg_cvd=coxph(Surv(time, cvd_death) ~ urxuhg, data=train, x=T)

summary(cox_uhg_cvd)

cox_bcd_cvd=coxph(Surv(time, cvd_death) ~ lbxbcd, data=train, x=T)

summary(cox_bcd_cvd)

cox_uur_cvd=coxph(Surv(time, cvd_death) ~ urxuur, data=train, x=T)

summary(cox_uur_cvd)

cox_udma_cvd=coxph(Surv(time, cvd_death) ~ urxudma, data=train, x=T)

summary(cox_udma_cvd)

cox_utu_cvd=coxph(Surv(time, cvd_death) ~ urxutu, data=train, x=T)

summary(cox_utu_cvd)

cox_usb_cvd=coxph(Surv(time, cvd_death) ~ urxusb, data=train, x=T)

summary(cox_usb_cvd)

cox_uab_cvd=coxph(Surv(time, cvd_death) ~ urxuab, data=train, x=T)

summary(cox_uab_cvd)

```

### Table

```{r}
HR_cvd=round(exp(c(cox_age_cvd$coefficients, cox_race_cvd$coefficients, cox_sex_cvd$coefficients, cox_smoke_cvd$coefficients, cox_sbp_cvd$coefficients, cox_totchol_cvd$coefficients, cox_hdl_cvd$coefficients, cox_bmi_cvd$coefficients, cox_hypertension_cvd$coefficients, cox_diabetes_cvd$coefficients)), 2)


Std_Err_cvd=round(c(summary(cox_age_cvd)$coefficients[3], unname(summary(cox_race_cvd)$coefficients[,3]), summary(cox_sex_cvd)$coefficients[3], summary(cox_smoke_cvd)$coefficients[3], summary(cox_sbp_cvd)$coefficients[3], summary(cox_totchol_cvd)$coefficients[3], summary(cox_hdl_cvd)$coefficients[3], summary(cox_bmi_cvd)$coefficients[,3], summary(cox_hypertension_cvd)$coefficients[3], summary(cox_diabetes_cvd)$coefficients[3]), 2)

confint_lb=round(c(summary(cox_age_cvd)$conf.int[3], unname(summary(cox_race_cvd)$conf.int[,3]), summary(cox_sex_cvd)$conf.int[3], summary(cox_smoke_cvd)$conf.int[3], summary(cox_sbp_cvd)$conf.int[3], summary(cox_totchol_cvd)$conf.int[3], summary(cox_hdl_cvd)$conf.int[3], unname(summary(cox_bmi_cvd)$conf.int[,3]), summary(cox_hypertension_cvd)$conf.int[3], summary(cox_diabetes_cvd)$conf.int[3]), 2)

confint_ub=round(c(summary(cox_age_cvd)$conf.int[4], unname(summary(cox_race_cvd)$conf.int[,4]), summary(cox_sex_cvd)$conf.int[4], summary(cox_smoke_cvd)$conf.int[4], summary(cox_sbp_cvd)$conf.int[4], summary(cox_totchol_cvd)$conf.int[4], summary(cox_hdl_cvd)$conf.int[4], unname(summary(cox_bmi_cvd)$conf.int[,4]), summary(cox_hypertension_cvd)$conf.int[4], summary(cox_diabetes_cvd)$conf.int[4]), 2)

p_vals_cvd=round(c(summary(cox_age_cvd)$coefficients[5], unname(summary(cox_race_cvd)$coefficients[,5]), summary(cox_sex_cvd)$coefficients[5], summary(cox_smoke_cvd)$coefficients[5], summary(cox_sbp_cvd)$coefficients[5], summary(cox_totchol_cvd)$coefficients[5], summary(cox_hdl_cvd)$coefficients[5], summary(cox_bmi_cvd)$coefficients[,5], summary(cox_hypertension_cvd)$coefficients[5], summary(cox_diabetes_cvd)$coefficients[5]), 3)

df_cvd=data.frame(HR_cvd, paste(confint_lb, ", ", confint_ub, sep=""), p_vals_cvd)

colnames(df_cvd)=c("HR", "95% CI", "p-value")

formattable(df_cvd,
    align=c("c", "c", "c"),
    list('p-value' = formatter(
      "span", style = x~style(font.weight=ifelse(x<0.05, "bold", "normal")))
    )
)


#Metals


HR_metals=round(exp(c(cox_ucs_cvd$coefficients, cox_umo_cvd$coefficients, cox_bpb_cvd$coefficients, cox_utl_cvd$coefficients, cox_uas_cvd$coefficients, cox_uco_cvd$coefficients, cox_uba_cvd$coefficients, cox_upb_cvd$coefficients, cox_ucd_cvd$coefficients, cox_thg_cvd$coefficients, cox_uhg_cvd$coefficients, cox_bcd_cvd$coefficients, cox_uur_cvd$coefficients, cox_udma_cvd$coefficients, cox_utu_cvd$coefficients, cox_usb_cvd$coefficients, cox_uab_cvd$coefficients)), 2)

Std_Err_metals=round(c(summary(cox_ucs_cvd)$coefficients[3], unname(summary(cox_umo_cvd)$coefficients[3]), summary(cox_bpb_cvd)$coefficients[3], summary(cox_utl_cvd)$coefficients[3], summary(cox_uas_cvd)$coefficients[3], summary(cox_uco_cvd)$coefficients[3], summary(cox_uba_cvd)$coefficients[3], summary(cox_upb_cvd)$coefficients[3], summary(cox_ucd_cvd)$coefficients[3], summary(cox_thg_cvd)$coefficients[3], summary(cox_uhg_cvd)$coefficients[3], summary(cox_bcd_cvd)$coefficients[3], summary(cox_uur_cvd)$coefficients[3], summary(cox_udma_cvd)$coefficients[3], summary(cox_utu_cvd)$coefficients[3], summary(cox_usb_cvd)$coefficients[3], summary(cox_uab_cvd)$coefficients[3]), 2)

confint_lb_metals=round(c(summary(cox_ucs_cvd)$conf.int[3], unname(summary(cox_umo_cvd)$conf.int[3]), summary(cox_bpb_cvd)$conf.int[3], summary(cox_utl_cvd)$conf.int[3], summary(cox_uas_cvd)$conf.int[3], summary(cox_uco_cvd)$conf.int[3], summary(cox_uba_cvd)$conf.int[3], unname(summary(cox_upb_cvd)$conf.int[3]), summary(cox_ucd_cvd)$conf.int[3], summary(cox_thg_cvd)$conf.int[3], summary(cox_uhg_cvd)$conf.int[3], summary(cox_bcd_cvd)$conf.int[3], summary(cox_uur_cvd)$conf.int[3], summary(cox_udma_cvd)$conf.int[3], unname(summary(cox_utu_cvd)$conf.int[3]), summary(cox_usb_cvd)$conf.int[3], summary(cox_uab_cvd)$conf.int[3]), 2)

confint_ub_metals=round(c(summary(cox_ucs_cvd)$conf.int[4], unname(summary(cox_umo_cvd)$conf.int[4]), summary(cox_bpb_cvd)$conf.int[4], summary(cox_utl_cvd)$conf.int[4], summary(cox_uas_cvd)$conf.int[4], summary(cox_uco_cvd)$conf.int[4], summary(cox_uba_cvd)$conf.int[4], unname(summary(cox_upb_cvd)$conf.int[4]), summary(cox_ucd_cvd)$conf.int[4], summary(cox_thg_cvd)$conf.int[4], summary(cox_uhg_cvd)$conf.int[4], summary(cox_bcd_cvd)$conf.int[4], summary(cox_uur_cvd)$conf.int[4], summary(cox_udma_cvd)$conf.int[4], unname(summary(cox_utu_cvd)$conf.int[4]), summary(cox_usb_cvd)$conf.int[4], summary(cox_uab_cvd)$conf.int[4]), 2)

p_vals_metals=round(c(summary(cox_ucs_cvd)$coefficients[5], unname(summary(cox_umo_cvd)$coefficients[5]), summary(cox_bpb_cvd)$coefficients[5], summary(cox_utl_cvd)$coefficients[5], summary(cox_uas_cvd)$coefficients[5], summary(cox_uco_cvd)$coefficients[5], summary(cox_uba_cvd)$coefficients[5], summary(cox_upb_cvd)$coefficients[5], summary(cox_ucd_cvd)$coefficients[5], summary(cox_thg_cvd)$coefficients[5], summary(cox_uhg_cvd)$coefficients[5], summary(cox_bcd_cvd)$coefficients[5], summary(cox_uur_cvd)$coefficients[5], summary(cox_udma_cvd)$coefficients[5], summary(cox_utu_cvd)$coefficients[5], summary(cox_usb_cvd)$coefficients[5], summary(cox_uab_cvd)$coefficients[5]), 3)



df_metals=data.frame(HR_metals, paste(confint_lb_metals, ", ", confint_ub_metals, sep=""), p_vals_metals)

colnames(df_metals)=c("HR", "95% CI", "p-value")

formattable(df_metals,
    align=c("c", "c", "c"),
    list('p-value' = formatter(
      "span", style = x~style(font.weight=ifelse(x<0.05, "bold", "normal")))
    )
)

```

## NRI

```{r}
#Traditional only
cox_survfit_trad<-survfit(CPHM_traditional_cvddeath,  newdata = test
                          )
cox_risk10_trad<-1-summary(cox_survfit_trad, times=10)$surv

#Traditional + blood metals

cox_survfit_trad_blood<-survfit(CPHM_blood_metals_main,  newdata = test)
cox_risk10_trad_blood<-1-summary(cox_survfit_trad_blood, times=10)$surv

#NRI
nri_cox_blood = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_blood, updown = "diff" , cut = .05)

nri_cox_blood$nri

# NRI=0.10 (0.05, 0.15)

#Continuous NRI

cont_nri_cox_trad_blood = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_blood, updown = "diff" , cut = 0) 

cont_nri_cox_trad_blood$nri

#Continuous NRI=0.23 (0.09, 0.38)


# Traditional + all metals

cox_survfit_trad_metals<-survfit(CPHM_all_metals_main,  newdata = test)
cox_risk10_trad_metals<-1-summary(cox_survfit_trad_metals, times=10)$surv

#NRI
nri_cox_trad = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_metals, updown = "diff" , cut = .05)

nri_cox_trad$nri

# NRI=0.080

#Continuous NRI

cont_nri_cox_trad = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_metals, updown = "diff" , cut = 0) 

cont_nri_cox_trad$nri

# NRI=0.208


#Traditional + metals + squared/interaction
cox_survfit_trad_int<-survfit(CPHM_all_metals_int,  newdata = test)

cox_risk10_trad_int<-1-summary(cox_survfit_trad_int, times=10)$surv

#NRI
nri_cox_int = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_int, updown = "diff" , cut = .05) 

nri_cox_int$nri # NRI=0.04 (-0.07, 0.14)

#Continuous NRI
cont_nri_cox_int = nribin(event=test$cvd_death, p.std = cox_risk10_trad, p.new = cox_risk10_trad_int, updown = "diff" , cut = 0) 

cont_nri_cox_int$nri # NRI=0.35 (0.21, 0.48)

```

##Partial R^2
```{r}



#Partial R^2 for age
partialr2_age = unname(coxr2(coxph(Surv(time, cvd_death) ~ age, data = train))$rsq)

#Adding one variable at a time to get the partial R^2 for each variable
partialr2_race = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace, data = train))$rsq) - partialr2_age

partialr2_sex = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male, data = train))$rsq) - partialr2_age - partialr2_race

partialr2_smk = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex

partialr2_sbp = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk

partialr2_tc = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk - partialr2_sbp

partialr2_hdl = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk - partialr2_sbp - partialr2_tc

partialr2_bmi = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk - partialr2_sbp - partialr2_tc - partialr2_hdl

partialr2_hypertension = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk - partialr2_sbp - partialr2_tc - partialr2_hdl - partialr2_bmi

partialr2_dm = unname(coxr2(coxph(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm, data = train))$rsq) - partialr2_age - partialr2_race - partialr2_sex - partialr2_smk - partialr2_sbp - partialr2_tc - partialr2_hdl - partialr2_bmi - partialr2_hypertension



#Confirming the sum of the partial R^2 is equal to the full model R^2
sum(partialr2_age, partialr2_bmi, partialr2_dm, partialr2_hdl, partialr2_hypertension, partialr2_race, partialr2_sbp, partialr2_sex, partialr2_smk, partialr2_tc)

unname(coxr2(CPHM_traditional_cvddeath)$rsq)



#Creating table

partialr2 = matrix(nrow = 10, ncol = 2)

partialr2[,1] = c("age", "race", "sex", "smk", "sbp", "tc", "hdl", "bmi", "hypertension", "dm")

partialr2[1,2] = partialr2_age
partialr2[2,2] = partialr2_race
partialr2[3,2] = partialr2_sex
partialr2[4,2] = partialr2_smk
partialr2[5,2] = partialr2_sbp
partialr2[6,2] = partialr2_tc
partialr2[7,2] = partialr2_hdl
partialr2[8,2] = partialr2_bmi
partialr2[9,2] = partialr2_hypertension
partialr2[10,2] = partialr2_dm

partialr2
```

# Elastic Net

## Pre-Processing

```{r}
###We need to change our data frames into matrices for the glmnet function
train_x_traditional = train %>%
  select(-c("all_death", "cvd_death", "time", metals$chemicals, "urxucr", "deathstat", "race", "bmi")) %>%
  data.matrix()

test_x_traditional = test %>%
  select(-c("all_death", "cvd_death", "time", metals$chemicals, "urxucr", "deathstat", "race", "bmi")) %>%
  data.matrix()

train_x_blood = train %>%
  select(-c("all_death", "cvd_death", "time", metals$chemicals[substr(metals$chemicals, 1, 1) == "u"], "urxucr", "deathstat", "race", "bmi")) %>%
  data.matrix()

test_x_blood = test %>%
  select(-c("all_death", "cvd_death", "time", metals$chemicals[substr(metals$chemicals, 1, 1) == "u"], "urxucr", "deathstat", "race", "bmi")) %>%
  data.matrix()


train_x_metals_main = train %>%
  select(-c("all_death", "cvd_death", "time", "deathstat", "race", "bmi")) %>%
  data.matrix()

test_x_metals_main=test %>%
  select(-c("all_death", "cvd_death", "time", "deathstat", "race", "bmi")) %>%
  data.matrix()

#The y outcome matrices from training data. glmnet function requires time and outcome as the "y" argument
y_cvd=train %>%
  select(c("time", "cvd_death")) %>%
  rename("status" = "cvd_death") %>%
  data.matrix()



#Chemicals
chemicals <- c(metals$chemicals, "urxucr")


chem_train=train[,chemicals]
chem_test=test[,chemicals]


#Create square terms of chemicals
square_chem_train<-data.frame(matrix(0,nrow(chem_train),ncol(chem_train)))
square_chem_test<-data.frame(matrix(0,nrow(chem_test),ncol(chem_test)))



for(i in 1:ncol(chem_train)) {
  square_chem_train[, i] <- chem_train[, i]^2
  colnames(square_chem_train)[i] <- paste(colnames(chem_train)[i], "^2", sep = "")
}

names_chem_sq_train<-colnames(square_chem_train)


for(i in 1:ncol(chem_test)) {
  square_chem_test[, i] <- chem_test[, i]^2
  colnames(square_chem_test)[i] <- paste(colnames(chem_test)[i], "^2", sep = "")
}

names_chem_sq_test<-colnames(square_chem_test)

#Create pair-wise interaction
pairs_chem_train <- data.frame(matrix(0, nrow(chem_train), choose(ncol(chem_train),2)))
pairs_chem_test <- data.frame(matrix(0, nrow(chem_test), choose(ncol(chem_test),2)))

k = 1
for(i in 1:(ncol(chem_train)-1)) {
  for(j in (i+1):ncol(chem_train)) {
    pairs_chem_train[, k] <- chem_train[, i]*chem_train[, j]
    colnames(pairs_chem_train)[k] <- paste(colnames(chem_train)[i], ".", colnames(chem_train)[j], sep = "")
    k = k + 1
  }
}

names_chem_int_train<-colnames(pairs_chem_train)

k = 1
for(i in 1:(ncol(chem_test)-1)) {
  for(j in (i+1):ncol(chem_test)) {
    pairs_chem_test[, k] <- chem_test[, i]*chem_test[, j]
    colnames(pairs_chem_test)[k] <- paste(colnames(chem_test)[i], ".", colnames(chem_test)[j], sep = "")
    k = k + 1
  }
}

names_chem_int_test<-colnames(pairs_chem_test)


train_x_metals_int=cbind(train_x_metals_main, square_chem_train, pairs_chem_train)

train_x_metals_int=data.matrix(train_x_metals_int)

test_x_metals_int=cbind(test_x_metals_main, square_chem_test, pairs_chem_test)

test_x_metals_int=data.matrix(test_x_metals_int)
```

## Traditional Variables

```{r}
###Use cross-validation to select optimal lambda
set.seed(56765)
cv_cvd_trad=cv.glmnet(train_x_traditional, y_cvd, family="cox", type.measure = "C", penalty.factor=c(1, rep(0, 4), rep(1, 5), rep(0, 3), rep(1, 2)))
plot(cv_cvd_trad)


fit_cvd_trad=list()

i=0
for(j in 1:11){
fit_cvd_trad[[j]]=glmnet(train_x_traditional, y_cvd, family="cox", alpha = i, lambda = 0.000510, penalty.factor=c(1, rep(0, 4), rep(1, 5), rep(0, 3), rep(1, 2))) #Penalty factor to ensure some levels of a categorical varible don't get shrunk to 0
i=i+0.1
}

###Predictions on Testing data

testing_trad=test %>%
  select(-c("all_death", "cvd_death", "time", metals$chemicals, "urxucr", "deathstat", "race", "bmi")) %>%
  data.matrix()

test_y_cvd=test %>%
  select(c("time", "cvd_death")) %>%
  rename("status" = "cvd_death") %>%
  data.matrix()


pred_cvd_trad=list()
cindex_cvd_trad=list()
cindex_cvd_trad_vec=vector()
for(i in 1:11){
pred_cvd_trad[[i]]=predict(fit_cvd_trad[[i]], newx = testing_trad, type="response", s = 0.000510)


cindex_cvd_trad[[i]]=Cindex(pred_cvd_trad[[i]], test_y_cvd)


cindex_cvd_trad_vec[i]=cindex_cvd_trad[[i]]
}
#Maximum c-index for each outcome

cindex_enet_trad=max(cindex_cvd_trad_vec) #0.8279
cindex_enet_trad

#Variables selected for optimal model

coef(fit_cvd_trad[[1]])
```

## Adding Blood Metals

```{r}
###Use cross-validation to select optimal lambda
set.seed(56765)
cv_cvd_blood=cv.glmnet(train_x_blood, y_cvd, family="cox", type.measure = "C", penalty.factor=c(rep(0, 15), rep(1, 3))) #Ensuring only metals can get shrunk to 0

plot(cv_cvd_blood)


fit_cvd_blood=list()

i=0
for(j in 1:11){
fit_cvd_blood[[j]]=glmnet(train_x_blood, y_cvd, family="cox", alpha = i, lambda = 0.0004828, penalty.factor=c(rep(0, 15), rep(1, 3)))
i=i+0.1
}

###Predictions on Testing data

pred_cvd_blood_main=list()
cindex_cvd_blood_main=list()
cindex_cvd_blood_vec_main=vector()
for(i in 1:11){

pred_cvd_blood_main[[i]]=predict(fit_cvd_blood[[i]], newx = test_x_blood, type="response", s=0.0004828)


cindex_cvd_blood_main[[i]]=Cindex(pred_cvd_blood_main[[i]], test_y_cvd)


cindex_cvd_blood_vec_main[i]=cindex_cvd_blood_main[[i]]
}

#Maximum c-index for each outcome

cindex_enet_blood=max(cindex_cvd_blood_vec_main) #0.8279
cindex_enet_blood

#Variables selected for optimal model

coef(fit_cvd_blood[[1]])
```


## Including all Metals

```{r}

#Only main effects
###Use cross-validation to select optimal lambda

set.seed(56765)
cv_cvd_metals_main=cv.glmnet(train_x_metals_main, y_cvd, family="cox", type.measure = "C", penalty.factor=c(rep(0, 15), rep(1, 18)))
plot(cv_cvd_metals_main)


fit_cvd_metals_main=list()

i=0
for(j in 1:11){
  fit_cvd_metals_main[[j]]=glmnet(train_x_metals_main, y_cvd, family="cox", lambda=0.001734, penalty.factor=c(rep(0, 15), rep(1, 18)), alpha = i)
i=i+0.1
}


###Predictions on Testing data



pred_cvd_metals_main=list()
cindex_cvd_metals_main=list()
cindex_cvd_metals_vec_main=vector()
for(i in 1:11){

pred_cvd_metals_main[[i]]=predict(fit_cvd_metals_main[[i]], newx = test_x_metals_main, type="response", s=0.001734)


cindex_cvd_metals_main[[i]]=Cindex(pred_cvd_metals_main[[i]], test_y_cvd)


cindex_cvd_metals_vec_main[i]=cindex_cvd_metals_main[[i]]
}

#Maximum c-index for each outcome

cindex_enet_metals_main=max(cindex_cvd_metals_vec_main) #0.8304 stand, 0.8320 not stand



#Variables selected for optimal model

coef(fit_cvd_metals_main[[11]])



#Including squared and interaction terms

#Exclude blood metals*urinary creatinine interactions

train_x_metals_int=subset(train_x_metals_int, select=-c(lbxbpb.urxucr, lbxthg.urxucr, lbxbcd.urxucr))
test_x_metals_int=subset(test_x_metals_int, select=-c(lbxbpb.urxucr, lbxthg.urxucr, lbxbcd.urxucr))

###Use cross-validation to select optimal lambda

set.seed(56765)
cv_cvd_metals_int=cv.glmnet(train_x_metals_int, y_cvd, family="cox", type.measure = "C", penalty.factor=c(rep(0, 15), rep(1, 186)))
plot(cv_cvd_metals_int)


fit_cvd_metals_int=list()

i=0
for(j in 1:11){
  fit_cvd_metals_int[[j]]=glmnet(train_x_metals_int, y_cvd, family="cox", lambda=0.004041, penalty.factor=c(rep(0, 15), rep(1, 186)), alpha = i)
i=i+0.1
}


###Predictions on Testing data



pred_cvd_metals_int=list()
cindex_cvd_metals_int=list()
cindex_cvd_metals_vec_int=vector()
for(i in 1:11){

pred_cvd_metals_int[[i]]=predict(fit_cvd_metals_int[[i]], newx = test_x_metals_int, type="response", s=0.004041)


cindex_cvd_metals_int[[i]]=Cindex(pred_cvd_metals_int[[i]], test_y_cvd)


cindex_cvd_metals_vec_int[i]=cindex_cvd_metals_int[[i]]
}

#Maximum c-index for each outcome

cindex_enet_metals_int=max(cindex_cvd_metals_vec_int) #0.8322 stand, 0.8249 not stand



#Variables selected for optimal model


coef(fit_cvd_metals_int[[8]])

```

## Comparison Table

```{r}

comparison_enet=matrix(nrow=4, ncol=1)
comparison_enet[1,]=c(cindex_enet_trad)
comparison_enet[2,]=c(cindex_enet_blood)
comparison_enet[3,]=c(cindex_enet_metals_main)
comparison_enet[4,]=c(cindex_enet_metals_int)


colnames(comparison_enet)=c("C-index")
rownames(comparison_enet)=c("Traditional Vars", "+ Blood Metals", "+ Urinary Metals", "+ Metals squared and interaction terms")
comparison_enet

```

## NRI

```{r, eval = F}
#Traditional only
enet_survfit_trad<-survfit(fit_cvd_trad[[1]], s = 0.000510, x = train_x_traditional, y = Surv(train$time, train$cvd_death), newx = test_x_traditional)


enet_risk10_trad<-1-summary(enet_survfit_trad, times=10)$surv



# Traditional + blood metals

enet_survfit_trad_blood<-survfit(fit_cvd_blood[[1]],  s = 0.0004828, x = train_x_blood, y = Surv(train$time, train$cvd_death), newx = test_x_blood)


enet_risk10_trad_blood<-1-summary(enet_survfit_trad_blood, times=10)$surv

#NRI
nri_enet_blood = nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_trad_blood, updown = "diff" , cut = 0.05)
nri_enet_blood$nri #NRI = 0.10 (0.05, 0.16)

#Continuous NRI
cont_nri_enet_blood=nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_trad_blood, updown = "diff" , cut = 0)
cont_nri_enet_blood$nri # NRI = 0.46 (0.32, 0.59)


# Traditional + metals

enet_survfit_trad_metals<-survfit(fit_cvd_metals_main[[11]],  s = 0.001734, x = train_x_metals_main, y = Surv(train$time, train$cvd_death), newx = test_x_metals_main)


enet_risk10_trad_metals<-1-summary(enet_survfit_trad_metals, times=10)$surv

#NRI
nri_enet_metals = nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_trad_metals, updown = "diff" , cut = 0.05)
nri_enet_metals$nri #NRI = 0.06 (0.02, 0.10)

#Continuous NRI
cont_nri_enet_metals=nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_trad_metals, updown = "diff" , cut = 0)
cont_nri_enet_metals$nri # NRI = 0.43 (0.29, 0.57)


#Traditional + metals + squared/interaction
enet_survfit_metals_int<-survfit(fit_cvd_metals_int[[8]],  s = 0.004041, x = train_x_metals_int, y = Surv(train$time, train$cvd_death), newx = test_x_metals_int)

enet_risk10_metals_int<-1-summary(enet_survfit_metals_int, times=10)$surv

#NRI
nri_enet_metals_int = nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_metals_int, updown = "diff" , cut = .05)
nri_enet_metals_int$nri #NRI = 0.05 (0.00, 0.10)

#Continuous NRI
cont_nri_enet_metals_int = nribin(event=test$cvd_death, p.std = enet_risk10_trad, p.new = enet_risk10_metals_int, updown = "diff" , cut = 0)
cont_nri_enet_metals_int$nri #NRI = 0.35 (0.20, 0.48)



```

# Random Forest

## Traditional Variables

```{r}
#Start by tuning ntree hyperparameter

rfcvdtrad_ntree=list()
predrfcvdtrad_ntree=list()
cindexcvdtrad_ntree=list()
for(ntree in c(100, 250, 500, 750, 1000, 1250, 1500)){
  index=toString(ntree)
  set.seed(56765)
  rfcvdtrad_ntree[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=ntree)
  
  
  predrfcvdtrad_ntree[[index]]=predict.rfsrc(rfcvdtrad_ntree[[index]], test)
  
  cindexcvdtrad_ntree[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdtrad_ntree[[index]]$predicted)
}

cindexcvdtrad_ntree=unlist(cindexcvdtrad_ntree)

cindexcvdtrad_ntree


#Since 250 trees has best performance, we will use that when experimenting with mtry (0.8328)
rfcvdtrad_mtry=list()
predrfcvdtrad_mtry=list()
cindexcvdtrad_mtry=list()

for(mtry in c(1:10)){
  index=toString(mtry)
  set.seed(56765)
  rfcvdtrad_mtry[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=250, mtry=mtry)
  
  
  predrfcvdtrad_mtry[[index]]=predict.rfsrc(rfcvdtrad_mtry[[index]], test)
  
  cindexcvdtrad_mtry[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdtrad_mtry[[index]]$predicted)
}

cindexcvdtrad_mtry=unlist(cindexcvdtrad_mtry)

cindexcvdtrad_mtry

#5 is optimal (0.8329)

###Now experimenting with nodesize using optimal ntree and mtry

rfcvdtrad_nodesize=list()
predrfcvdtrad_nodesize=list()
cindexcvdtrad_nodesize=list()

for(nodesize in c(1:25)){
  index=toString(nodesize)
  set.seed(56765)
  rfcvdtrad_nodesize[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=250, mtry=5, nodesize=nodesize)
  
  
  predrfcvdtrad_nodesize[[index]]=predict.rfsrc(rfcvdtrad_nodesize[[index]], test)
  
  cindexcvdtrad_nodesize[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdtrad_nodesize[[index]]$predicted)
  print(nodesize)
}

cindexcvdtrad_nodesize=unlist(cindexcvdtrad_nodesize)

cindexcvdtrad_nodesize

##24 is optimal (0.8345)

#Finally, tuning nsplit

rfcvdtrad_nsplit=list()
predrfcvdtrad_nsplit=list()
cindexcvdtrad_nsplit=list()

for(nsplit in c(1:20)){
  index=toString(nsplit)
  set.seed(56765)
  rfcvdtrad_nsplit[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=250, mtry=5, nodesize=24, nsplit=nsplit)
  
  
  predrfcvdtrad_nsplit[[index]]=predict.rfsrc(rfcvdtrad_nsplit[[index]], test)
  
  cindexcvdtrad_nsplit[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdtrad_nsplit[[index]]$predicted)
}

cindexcvdtrad_nsplit=unlist(cindexcvdtrad_nsplit)

cindexcvdtrad_nsplit

#Try nsplit=0?

set.seed(56765)
rfcvdtrad_nsplit0=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=250, mtry=5, nodesize=23, nsplit=0)

predrfcvdtrad_nsplit0=predict.rfsrc(rfcvdtrad_nsplit0, test)
  
cindexcvdtrad_nsplit0=1-get.cindex(test$time, test$cvd_death, predrfcvdtrad_nsplit0$predicted)

cindexcvdtrad_nsplit0

#5 is optimal (0.8353)

#Now the hyperparameters have been tuned, let's create the optimal forest
set.seed(56765)
optimalrf_cvd_trad=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm, data=train, ntree=250, mtry=5, nodesize=24, nsplit=5)

pred_optimalrf_cvd_trad=predict.rfsrc(optimalrf_cvd_trad, test)

rf_cindex_cvd_trad=1-get.cindex(test$time, test$cvd_death, pred_optimalrf_cvd_trad$predicted)
```

## Including Blood Metals

```{r}

#Start by tuning ntree
rfcvdblood_ntree=list()
predrfcvdblood_ntree=list()
cindexcvdblood_ntree=list()

for(ntree in c(100, 250, 500, 750, 1000, 1250, 1500)){
  index=toString(ntree)
  set.seed(56765)
  rfcvdblood_ntree[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=ntree)
  
  predrfcvdblood_ntree[[index]]=predict.rfsrc(rfcvdblood_ntree[[index]], test)
  
  cindexcvdblood_ntree[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdblood_ntree[[index]]$predicted)
}

cindexcvdblood_ntree=unlist(cindexcvdblood_ntree)

cindexcvdblood_ntree


#Since 500 trees has best performance (0.8288), we will use that when experimenting with mtry
rfcvdblood_mtry=list()
predrfcvdblood_mtry=list()
cindexcvdblood_mtry=list()

for(mtry in c(1:15)){
  index=toString(mtry)
  set.seed(56765)
  rfcvdblood_mtry[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=500, mtry = mtry)
  
  
  predrfcvdblood_mtry[[index]]=predict.rfsrc(rfcvdblood_mtry[[index]], test)
  
  cindexcvdblood_mtry[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdblood_mtry[[index]]$predicted)
}

cindexcvdblood_mtry=unlist(cindexcvdblood_mtry)

cindexcvdblood_mtry

#3 is optimal (0.8292)

###Now experimenting with nodesize using optimal ntree and mtry

rfcvdblood_nodesize=list()
predrfcvdblood_nodesize=list()
cindexcvdblood_nodesize=list()

for(nodesize in c(1:20)){
  index=toString(nodesize)
  set.seed(56765)
  rfcvdblood_nodesize[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=500, mtry=3, nodesize=nodesize)
  
  
  predrfcvdblood_nodesize[[index]]=predict.rfsrc(rfcvdblood_nodesize[[index]], test)
  
  cindexcvdblood_nodesize[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdblood_nodesize[[index]]$predicted)
}

cindexcvdblood_nodesize=unlist(cindexcvdblood_nodesize)

cindexcvdblood_nodesize

##15 is optimal(0.8292)

#Finally, tuning nsplit

rfcvdblood_nsplit=list()
predrfcvdblood_nsplit=list()
cindexcvdblood_nsplit=list()

for(nsplit in c(1:20)){
  index=toString(nsplit)
  set.seed(56765)
  rfcvdblood_nsplit[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=500, mtry=3, nodesize=15, nsplit=nsplit)
  
  predrfcvdblood_nsplit[[index]]=predict.rfsrc(rfcvdblood_nsplit[[index]], test)
  
  cindexcvdblood_nsplit[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdblood_nsplit[[index]]$predicted)
}

cindexcvdblood_nsplit=unlist(cindexcvdblood_nsplit)

cindexcvdblood_nsplit

#Try nsplit=0?
set.seed(56765)
rfcvdblood_nsplit0=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=500, mtry=3, nodesize=15, nsplit=0)

predrfcvdblood_nsplit0=predict.rfsrc(rfcvdblood_nsplit0, test)
  
cindexcvdblood_nsplit0=1-get.cindex(test$time, test$cvd_death, predrfcvdblood_nsplit0$predicted)

cindexcvdblood_nsplit0

#5 is optimal (0.8317)

#Now the hyperparameters have been tuned, let's create the optimal forest
set.seed(56765)
optimalrf_cvd_blood=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + lbxbpb + lbxthg + lbxbcd, data=train, ntree=500, mtry=3, nodesize=15, nsplit=5)

pred_optimalrf_cvd_blood=predict.rfsrc(optimalrf_cvd_blood, test)

rf_cindex_cvd_blood=1-get.cindex(test$time, test$cvd_death, pred_optimalrf_cvd_blood$predicted)

```

## Including all Metals

```{r}

#Start by tuning ntree
rfcvdmetals_ntree=list()
predrfcvdmetals_ntree=list()
cindexcvdmetals_ntree=list()

for(ntree in c(100, 250, 500, 750, 1000, 1250, 1500)){
  index=toString(ntree)
  set.seed(56765)
  rfcvdmetals_ntree[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=ntree)
  
  predrfcvdmetals_ntree[[index]]=predict.rfsrc(rfcvdmetals_ntree[[index]], test)
  
  cindexcvdmetals_ntree[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdmetals_ntree[[index]]$predicted)
}

cindexcvdmetals_ntree=unlist(cindexcvdmetals_ntree)

cindexcvdmetals_ntree


#Since 100 trees has best performance (0.8241), we will use that when experimenting with mtry
rfcvdmetals_mtry=list()
predrfcvdmetals_mtry=list()
cindexcvdmetals_mtry=list()

for(mtry in c(1:15)){
  index=toString(mtry)
  set.seed(56765)
  rfcvdmetals_mtry[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=100, mtry=mtry)
  
  
  predrfcvdmetals_mtry[[index]]=predict.rfsrc(rfcvdmetals_mtry[[index]], test)
  
  cindexcvdmetals_mtry[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdmetals_mtry[[index]]$predicted)
}

cindexcvdmetals_mtry=unlist(cindexcvdmetals_mtry)

cindexcvdmetals_mtry

#6 is optimal (0.8241)

###Now experimenting with nodesize using optimal ntree and mtry

rfcvdmetals_nodesize=list()
predrfcvdmetals_nodesize=list()
cindexcvdmetals_nodesize=list()

for(nodesize in c(1:20)){
  index=toString(nodesize)
  set.seed(56765)
  rfcvdmetals_nodesize[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=100, mtry=6, nodesize=nodesize)
  
  
  predrfcvdmetals_nodesize[[index]]=predict.rfsrc(rfcvdmetals_nodesize[[index]], test)
  
  cindexcvdmetals_nodesize[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdmetals_nodesize[[index]]$predicted)
}

cindexcvdmetals_nodesize=unlist(cindexcvdmetals_nodesize)

cindexcvdmetals_nodesize

##14 is optimal(0.8262)

#Finally, tuning nsplit

rfcvdmetals_nsplit=list()
predrfcvdmetals_nsplit=list()
cindexcvdmetals_nsplit=list()

for(nsplit in c(1:20)){
  index=toString(nsplit)
  set.seed(56765)
  rfcvdmetals_nsplit[[index]]=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=100, mtry=6, nodesize=14, nsplit=nsplit)
  
  predrfcvdmetals_nsplit[[index]]=predict.rfsrc(rfcvdmetals_nsplit[[index]], test)
  
  cindexcvdmetals_nsplit[[index]]=1-get.cindex(test$time, test$cvd_death, predrfcvdmetals_nsplit[[index]]$predicted)
}

cindexcvdmetals_nsplit=unlist(cindexcvdmetals_nsplit)

cindexcvdmetals_nsplit

#Try nsplit=0?
set.seed(56765)
rfcvdmetals_nsplit0=rfsrc(Surv(time, cvd_death) ~ age + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=100, mtry=6, nodesize=14, nsplit=0)

predrfcvdmetals_nsplit0=predict.rfsrc(rfcvdmetals_nsplit0, test)
  
cindexcvdmetals_nsplit0=1-get.cindex(test$time, test$cvd_death, predrfcvdmetals_nsplit0$predicted)

cindexcvdmetals_nsplit0

#10 is optimal (0.8262)

#Now the hyperparameters have been tuned, let's create the optimal forest
set.seed(56765)
optimalrf_cvd_metals=rfsrc(Surv(time, cvd_death) ~ age + white + black + hispanic + otherrace + male + currentsmk + sbp + lbxtc + hdl + normal_weight + overweight + obese + hypertension + dm + urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data=train, ntree=100, mtry=6, nodesize=14, nsplit=10)

pred_optimalrf_cvd_metals=predict.rfsrc(optimalrf_cvd_metals, test)

rf_cindex_cvd_metals=1-get.cindex(test$time, test$cvd_death, pred_optimalrf_cvd_metals$predicted)
```

## Comparison Table

```{r}
comparison_rf=matrix(nrow=3, ncol=1)
comparison_rf[1,]=c(rf_cindex_cvd_trad)
comparison_rf[2,]=c(rf_cindex_cvd_blood)
comparison_rf[3,]=c(rf_cindex_cvd_metals)

colnames(comparison_rf)=c("C-index")
rownames(comparison_rf)=c("Traditional Vars", "+Blood Metals", "+Urinary Metals")

comparison_rf = rbind(comparison_rf, comparison_rf[3])
comparison_rf
```

## NRI
```{r}

#Traditional vars
rf_survfit_trad = pred_optimalrf_cvd_trad$survival[,which(pred_optimalrf_cvd_trad$time.interest==10)]

#Blood metals
rf_survfit_trad_blood=pred_optimalrf_cvd_metals$survival[,which(pred_optimalrf_cvd_blood$time.interest==10)]

#NRI
nri_rf_blood = nribin(event=test$cvd_death, p.std = 1-rf_survfit_trad, p.new = 1-rf_survfit_trad_blood, updown = "diff" , cut = 0.05) #NRI = -0.151

nri_rf_blood$nri #NRI = -0.15 (-0.23, -0.08)

#Continuous NRI
cont_nri_rf_blood = nribin(event=test$cvd_death, p.std = 1-rf_survfit_trad, p.new = 1-rf_survfit_trad_blood, updown = "diff" , cut = 0) 

cont_nri_rf_blood$nri #Continuous NRI = -0.49 (-0.63, -0.36)

#All metals
rf_survfit_trad_metals=pred_optimalrf_cvd_metals$survival[,which(pred_optimalrf_cvd_metals$time.interest==10)]

#NRI
nri_rf_metals = nribin(event=test$cvd_death, p.std = 1-rf_survfit_trad, p.new = 1-rf_survfit_trad_metals, updown = "diff" , cut = 0.05)

nri_rf_metals$nri  #NRI = -0.151 (-0.23, -0.08)

#Continuous NRI
cont_nri_rf_metals = nribin(event=test$cvd_death, p.std = 1-rf_survfit_trad, p.new = 1-rf_survfit_trad_metals, updown = "diff" , cut = 0) 

cont_nri_rf_metals$nri #NRI = -0.493 (-0.64, -0.35)
```

#Figure for Presentation

```{r}
comparison_df = data.frame(cbind(comparison_CPHM, comparison_enet, comparison_rf))

plot_df = comparison_df %>%
  rownames_to_column() %>%
  rename("CPHM" = C.index, "ENET" = C.index.1, "Survival_RF" = C.index.2, "Predictors" = rowname) %>%
  pivot_longer(cols = c(CPHM, ENET, Survival_RF), names_to = "Model", values_to = "C.index")

plot_df$Predictors = factor(plot_df$Predictors, levels = c("Traditional Vars", "+Blood Metals (main effects)", "+Urinary Metals (main effects)", "+ Metals squared and interaction"))

ggplot(data = plot_df, aes(x = Predictors, y = C.index, color = Model, group = Model)) + geom_line(size = 1) + geom_point(size = 5)  + labs(y = "C-index")  + theme_bw(base_size = 18) + scale_x_discrete(labels = c("Traditional Predictors", "+ Blood Metals", "+ Urinary Metals", "+ Metals squared and interaction")) + theme(axis.text = element_text(face = "bold")) + scale_color_discrete(name = "Model", labels = c("Cox Proportional Hazards Model", "Elastic-Net", "Survival Random Forest"))


```


# Framingham Categories

##Pre-processing
```{r}

#Risk scores for men (From Figure 3 of 1998 paper)
fram_train_women = train %>%
  filter(male == 0) %>%
  select(age, lbxtc, hdl, sbp, currentsmk, dm, male, time, cvd_death, metals$chemicals, urxucr) %>%
  mutate("age_cat" = ifelse(age >= 40 & age <= 44, 0, 
                            ifelse(age >=45 & age <=49, 3,
                                   ifelse(age >= 50 & age <= 54, 6,
                                          ifelse(age >= 55 & age <= 59, 7,
                                                 8)))),
         "tc_cat" = ifelse(lbxtc < 160, -2,
                           ifelse(lbxtc >= 160 & lbxtc <= 199, 0,
                                  ifelse(lbxtc >= 200 & lbxtc <= 239, 1,
                                         ifelse(lbxtc >= 240 & lbxtc <= 279, 1,
                                                3)))),
         "hdl_cat" = ifelse(hdl < 35, 5,
                            ifelse(hdl >=35 & hdl <= 44, 2,
                                   ifelse(hdl >= 45 & hdl <= 49, 1,
                                          ifelse(hdl >= 50 & hdl <= 59, 0,
                                                 -3)))),
         "sbp_cat" = ifelse(sbp < 120, -3,
                            ifelse(sbp >=120 & sbp <=139, 0,
                                   ifelse(sbp >= 140 & sbp <= 159, 2,
                                          3))),
         "currentsmk_cat" = ifelse(currentsmk == 1, 2, 0),
         "dm_cat" = ifelse(dm == 1, 4, 0)) %>%
  rownames_to_column() %>%
  group_by(rowname) %>%
  mutate("score" = sum(age_cat, tc_cat, hdl_cat, sbp_cat, currentsmk_cat, dm_cat))


#Risk scores for women (from Figure 4 of 1998 paper)
fram_train_men = train %>%
  filter(male == 1) %>%
  select(age, lbxtc, hdl, sbp, currentsmk, dm, male, time, cvd_death, metals$chemicals, urxucr) %>%
  mutate("age_cat" = ifelse(age >= 40 & age <= 44, 1, 
                            ifelse(age >=45 & age <=49, 2,
                                   ifelse(age >= 50 & age <= 54, 3,
                                          ifelse(age >= 55 & age <= 59, 4,
                                                 ifelse(age >= 60 & age <= 64, 5,
                                                        ifelse(age >= 65 & age <= 69, 6,
                                                               7)))))),
         "tc_cat" = ifelse(lbxtc < 160, -3,
                           ifelse(lbxtc >= 160 & lbxtc <= 199, 0,
                                  ifelse(lbxtc >= 200 & lbxtc <= 239, 1,
                                         ifelse(lbxtc >= 240 & lbxtc <= 279, 2,
                                                3)))),
         "hdl_cat" = ifelse(hdl < 35, 2,
                            ifelse(hdl >=35 & hdl <= 44, 1,
                                   ifelse(hdl >= 45 & hdl <= 49, 0,
                                          ifelse(hdl >= 50 & hdl <= 59, 0,
                                                 -2)))),
         "sbp_cat" = ifelse(sbp < 120, 0,
                            ifelse(sbp >=120 & sbp <=129, 0,
                                   ifelse(sbp >=130 & sbp <= 139, 1,
                                          ifelse(sbp >= 140 & sbp <= 159, 2,
                                          3)))),
         "currentsmk_cat" = ifelse(currentsmk == 1, 2, 0),
         "dm_cat" = ifelse(dm == 1, 2, 0)) %>%
  rownames_to_column() %>%
  group_by(rowname) %>%
  mutate("score" = sum(age_cat, tc_cat, hdl_cat, sbp_cat, currentsmk_cat, dm_cat))


fram_test_women = test %>%
  filter(male == 0) %>%
  select(age, lbxtc, hdl, sbp, currentsmk, dm, male, time, cvd_death, metals$chemicals, urxucr) %>%
  mutate("age_cat" = ifelse(age >= 40 & age <= 44, 0, 
                            ifelse(age >=45 & age <=49, 3,
                                   ifelse(age >= 50 & age <= 54, 6,
                                          ifelse(age >= 55 & age <= 59, 7,
                                                 8)))),
         "tc_cat" = ifelse(lbxtc < 160, -2,
                           ifelse(lbxtc >= 160 & lbxtc <= 199, 0,
                                  ifelse(lbxtc >= 200 & lbxtc <= 239, 1,
                                         ifelse(lbxtc >= 240 & lbxtc <= 279, 1,
                                                3)))),
         "hdl_cat" = ifelse(hdl < 35, 5,
                            ifelse(hdl >=35 & hdl <= 44, 2,
                                   ifelse(hdl >= 45 & hdl <= 49, 1,
                                          ifelse(hdl >= 50 & hdl <= 59, 0,
                                                 -3)))),
         "sbp_cat" = ifelse(sbp < 120, -3,
                            ifelse(sbp >=120 & sbp <=139, 0,
                                   ifelse(sbp >= 140 & sbp <= 159, 2,
                                          3))),
         "currentsmk_cat" = ifelse(currentsmk == 1, 2, 0),
         "dm_cat" = ifelse(dm == 1, 4, 0)) %>%
  rownames_to_column() %>%
  group_by(rowname) %>%
  mutate("score" = sum(age_cat, tc_cat, hdl_cat, sbp_cat, currentsmk_cat, dm_cat))


fram_test_men = test %>%
  filter(male == 1) %>%
  select(age, lbxtc, hdl, sbp, currentsmk, dm, male, time, cvd_death, metals$chemicals, urxucr) %>%
  mutate("age_cat" = ifelse(age >= 40 & age <= 44, 1, 
                            ifelse(age >=45 & age <=49, 2,
                                   ifelse(age >= 50 & age <= 54, 3,
                                          ifelse(age >= 55 & age <= 59, 4,
                                                 ifelse(age >= 60 & age <= 64, 5,
                                                        ifelse(age >= 65 & age <= 69, 6,
                                                               7)))))),
         "tc_cat" = ifelse(lbxtc < 160, -3,
                           ifelse(lbxtc >= 160 & lbxtc <= 199, 0,
                                  ifelse(lbxtc >= 200 & lbxtc <= 239, 1,
                                         ifelse(lbxtc >= 240 & lbxtc <= 279, 2,
                                                3)))),
         "hdl_cat" = ifelse(hdl < 35, 2,
                            ifelse(hdl >=35 & hdl <= 44, 1,
                                   ifelse(hdl >= 45 & hdl <= 49, 0,
                                          ifelse(hdl >= 50 & hdl <= 59, 0,
                                                 -2)))),
         "sbp_cat" = ifelse(sbp < 120, 0,
                            ifelse(sbp >=120 & sbp <=129, 0,
                                   ifelse(sbp >=130 & sbp <= 139, 1,
                                          ifelse(sbp >= 140 & sbp <= 159, 2,
                                          3)))),
         "currentsmk_cat" = ifelse(currentsmk == 1, 2, 0),
         "dm_cat" = ifelse(dm == 1, 2, 0)) %>%
  rownames_to_column() %>%
  group_by(rowname) %>%
  mutate("score" = sum(age_cat, tc_cat, hdl_cat, sbp_cat, currentsmk_cat, dm_cat))

```




##Descriptive Statistics
```{r}

#Combining datasets
fram = rbind(fram_train_men, fram_train_women, fram_test_men, fram_test_women)

#Creating mirror histogram
fram_plot = fram %>%
  mutate(score_male = ifelse(male == 1, score, NA),
         score_female = ifelse(male == 0, score, NA)) 

ggplot(data = fram_plot, aes(x = x)) + geom_histogram(aes(x = score_male, y = ..count..), color = "white", fill = "cyan4") + geom_label(aes(x = 15, y = 200, label = "Male"), color = "cyan4") + geom_histogram(aes(x = score_female, y = -..count..), color = "white", fill = "blue") + geom_label(aes(x = 15, y = -200, label = "Female"), color = "blue") + labs(title = "Histogram of Framingham CHD risk scores grouped by sex", x = "Score") + theme_gray()

#Getting summary statistics
fram_plot %>%
  group_by(male) %>%
  summarize(mean(score),
            sd(score),
            median(score),
            min(score),
            max(score))
```

## Cox Model only using score
```{r}

#Creating models
fram_cox_men = coxph(Surv(time, cvd_death) ~ score, x = T, data = fram_train_men)
summary(fram_cox_men)

fram_cox_women = coxph(Surv(time, cvd_death) ~ score, x = T, data = fram_train_women)
summary(fram_cox_women)

#Getting c-index for men
cindex_men = cindex(fram_cox_men, Surv(time, cvd_death) ~ score, data = fram_test_men)$AppCindex
cindex_men
#Getting c-index for women
cindex_women = cindex(fram_cox_women, Surv(time, cvd_death) ~ score, data = fram_test_women)$AppCindex
cindex_women

```

## Adding Blood Metals

```{r}

fram_cox_men_blood_metals = coxph(Surv(time, cvd_death) ~ score + lbxbpb + lbxthg + lbxbcd, x = T, data = fram_train_men)
summary(fram_cox_men_blood_metals)

fram_cox_women_blood_metals = coxph(Surv(time, cvd_death) ~ score + lbxbpb + lbxthg + lbxbcd, x = T, data = fram_train_women)
summary(fram_cox_women_blood_metals)


#Getting c-index for men
cindex_men_blood_metals = cindex(fram_cox_men_blood_metals, Surv(time, cvd_death) ~ score + lbxbpb + lbxthg + lbxbcd, data = fram_test_men)$AppCindex
cindex_men_blood_metals

#Getting c-index for women
cindex_women_blood_metals = cindex(fram_cox_women_blood_metals, Surv(time, cvd_death) ~ score + lbxbpb + lbxthg + lbxbcd, data = fram_test_women)$AppCindex
cindex_women_blood_metals
```

## Adding Urinary Metals
```{r}
#Creating models
fram_cox_men_all_metals = coxph(Surv(time, cvd_death) ~ score + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, x = T, data = fram_train_men)

summary(fram_cox_men_all_metals)

fram_cox_women_all_metals = coxph(Surv(time, cvd_death) ~ score + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, x = T, data = fram_train_women)

summary(fram_cox_women_all_metals)

#Getting c-index for men
cindex_men_all_metals = cindex(fram_cox_men_all_metals, Surv(time, cvd_death) ~ score + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data = fram_test_men)$AppCindex
cindex_men_all_metals

#Getting c-index for women
cindex_women_all_metals = cindex(fram_cox_women_all_metals, Surv(time, cvd_death) ~ score + urxucs + urxumo + lbxbpb + urxutl + urxuas + urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr, data = fram_test_women)$AppCindex
cindex_women_all_metals


```

## With squared and interaction terms
```{r}
fram_cox_men_int=coxph(Surv(time, cvd_death) ~ score + urxucs + urxucs^2 + urxumo + urxumo^2 + lbxbpb + lbxbpb^2 + urxutl + urxutl^2 + urxuas + urxuas^2 + urxuco + urxuco^2 + urxuba + urxuba^2 + urxupb + urxupb^2 + urxucd + urxucd^2 + lbxthg + lbxthg^2 + urxuhg + urxuhg^2 + lbxbcd + lbxbcd^2 + urxuur + urxuur^2 + urxudma + urxudma^2 + urxutu + urxutu^2 + urxusb + urxusb^2 + urxuab + urxuab^2 + urxucr + urxucr^2 + (urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr)^2  - lbxbpb:urxucr - lbxthg:urxucr - lbxbcd:urxucr, data=fram_train_men, x=T)

summary(fram_cox_men_int)

###WARNING: DID NOT CONVERGE Too many predictors (169) and not enough cases of CVD Mortality (119)

fram_cox_women_int=coxph(Surv(time, cvd_death) ~ score + urxucs + urxucs^2 + urxumo + urxumo^2 + lbxbpb + lbxbpb^2 + urxutl + urxutl^2 + urxuas + urxuas^2 + urxuco + urxuco^2 + urxuba + urxuba^2 + urxupb + urxupb^2 + urxucd + urxucd^2 + lbxthg + lbxthg^2 + urxuhg + urxuhg^2 + lbxbcd + lbxbcd^2 + urxuur + urxuur^2 + urxudma + urxudma^2 + urxutu + urxutu^2 + urxusb + urxusb^2 + urxuab + urxuab^2 + urxucr + urxucr^2 + (urxucs + urxumo + lbxbpb + urxutl + urxuas+ urxuco + urxuba + urxupb + urxucd + lbxthg + urxuhg + lbxbcd + urxuur + urxudma + urxutu + urxusb + urxuab + urxucr)^2  - lbxbpb:urxucr - lbxthg:urxucr - lbxbcd:urxucr, data=fram_train_women, x=T)

summary(fram_cox_women_int)

###WARNING: DID NOT CONVERGE (Too many predictors (169) and not enough cases of CVD Mortality (92))

#Getting c-index for men
cindex_men_int = cindex(fram_cox_men_int, Surv(time, cvd_death) ~ score, data = fram_test_men)$AppCindex
cindex_men_int

#Getting c-index for women
cindex_women_int = cindex(fram_cox_women_int, Surv(time, cvd_death) ~ score, data = fram_test_women)$AppCindex
cindex_women_int
```
## Comparison table
```{r}
comparison_fram=matrix(nrow=3, ncol=2)
comparison_fram[1,1]=unlist(unname(cindex_men))
comparison_fram[2,1]=unlist(unname(cindex_men_metals))
comparison_fram[3,1]=unlist(unname(cindex_men_int))
comparison_fram[1,2]=unlist(unname(cindex_women))
comparison_fram[2,2]=unlist(unname(cindex_women_metals))
comparison_fram[3,2]=unlist(unname(cindex_women_int))

colnames(comparison_fram)=c("Men", "Women")
rownames(comparison_fram)=c("Score", "+ metals", "+ metals squared and interaction")

comparison_fram
```
## NRI
### Men
```{r}
#Score only
fram_survfit<-survfit(fram_cox_men,  newdata = fram_test_men)

fram_risk10<-1-summary(fram_survfit, times=10)$surv



#Score + metals

fram_survfit_metals<-survfit(fram_cox_men_metals,  newdata = fram_test_men)
fram_risk10_metals<-1-summary(fram_survfit_metals, times=10)$surv

#NRI
nri_fram_metals = nribin(event=fram_test_men$cvd_death, p.std = fram_risk10, p.new = fram_risk10_metals, updown = "diff" , cut = .05) #NRI = -0.006

#Continuous NRI
cont_nri_fram_metals = nribin(event=fram_test_men$cvd_death, p.std = fram_risk10, p.new = fram_risk10_metals, updown = "diff" , cut = 0) # NRI=0.3288


#Score + metals + squared/interaction
fram_survfit_int<-survfit(fram_cox_men_int,  newdata = fram_test_men)
fram_risk10_int<-1-summary(fram_survfit_int, times=10)$surv

#NRI
nri_fram_int = nribin(event=fram_test_men$cvd_death, p.std = fram_risk10, p.new = fram_risk10_int, updown = "diff" , cut = .05) #NRI = -0.24

#Continuous NRI
cont_nri_cox_int = nribin(event=fram_test_men$cvd_death, p.std = fram_risk10, p.new = fram_risk10_int, updown = "diff" , cut = 0) #NRI = 0 (Didn't run correctly because the model didn't converge)
```

### Women
```{r}
#Score only
fram_survfit<-survfit(fram_cox_women,  newdata = fram_test_women)

fram_risk10<-1-summary(fram_survfit, times=10)$surv

#Score + metals

fram_survfit_metals<-survfit(fram_cox_women_metals,  newdata = fram_test_women)
fram_risk10_metals<-1-summary(fram_survfit_metals, times=10)$surv

#NRI
nri_fram_metals = nribin(event=fram_test_women$cvd_death, p.std = fram_risk10, p.new = fram_risk10_metals, updown = "diff" , cut = .05) #NRI = 0.050

#Continuous NRI
cont_nri_fram_metals = nribin(event=fram_test_women$cvd_death, p.std = fram_risk10, p.new = fram_risk10_metals, updown = "diff" , cut = 0) # NRI=0.1743


#Score + metals + squared/interaction
fram_survfit_int<-survfit(fram_cox_women_int,  newdata = fram_test_women)
fram_risk10_int<-1-summary(fram_survfit_int, times=10)$surv

#NRI
nri_fram_int = nribin(event=fram_test_women$cvd_death, p.std = fram_risk10, p.new = fram_risk10_int, updown = "diff" , cut = .05) #NRI = -0.017

#Continuous NRI
cont_nri_cox_int = nribin(event=fram_test_women$cvd_death, p.std = fram_risk10, p.new = fram_risk10_int, updown = "diff" , cut = 0) #NRI = 0.157
```